---
title: "Introduction to csstyle"
author: "Richard Aubrey White, Chi Zhang"
date: "2021-02-10"
output:
  rmarkdown::html_vignette:
    fig_width: 6
    fig_height: 6
vignette: >
  %\VignetteIndexEntry{Introduction to csstyle}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Checklist

- Make sure that your dataset contains all days/weeks (i.e. there can be days/weeks with 0 cases)
- Use `q <- q + csstyle::theme_cs()`
- Use `q <- q + geom_col(fill = csstyle::base_color, width = 0.8)`
- For `scale_y_continuous` use `breaks = csstyle::pretty_breaks(5)`
- For `scale_y_continuous` use `expand = expansion(mult = c(0, 0.1))`
- If the x-axis is too cramped, make the labels vertical `q <- q + csstyle::set_x_axis_vertical()`
- Say when your data was extracted `q <- q + labs(caption = "Data extracted on 2018-02-20")`

If multiple geographical areas:
- Use `q <- q + lemon::facet_rep_wrap(~location_code, repeat.tick.labels = "y")`

```{r}
library(ggplot2)
library(data.table)
```

## Graphs for one geographical area

Daily epicurve for `county_nor03`

```{r warning=FALSE}
csstyle::plot_epicurve(
  covidnor::total[
    granularity_geo=="nation" &
    granularity_time=="day" &
    isoyear==2020
  ],
  type = "single",
  var_x = "date",
  var_y = "hospital_admissions_main_cause_n"
)
```

Weekly epicurve for `county_nor03`

```{r}
# q <- ggplot(w[location_code=="county_nor03"], aes(x = isoyearweek, y = N))
# q <- q + geom_col(fill = csstyle::base_color, width = 0.8)
# q <- q + scale_x_discrete("Isoweek")
# q <- q + scale_y_continuous("Number of reported cases",
#   breaks = csstyle::pretty_breaks(5),
#   expand = expansion(mult = c(0, 0.1))
# )
# q <- q + labs(title = "Epicurve from 2018-01-01 to 2018-02-20")
# q <- q + labs(caption = fhi_caption())
# q <- q + csstyle::theme_fhi_lines_horizontal()
# q
```

Weekly epicurve with vertical x-axis labels

```{r}
# q <- ggplot(w, aes(x = isoyearweek, y = N))
# q <- q + geom_col(fill = csstyle::base_color, width = 0.8)
# q <- q + scale_x_discrete("Isoweek")
# q <- q + scale_y_continuous("Number of reported cases",
#   breaks = csstyle::pretty_breaks(5),
#   expand = expansion(mult = c(0, 0.1))
# )
# q <- q + labs(title = "Epicurve from 2018-01-01 to 2018-02-20")
# q <- q + labs(caption = fhi_caption())
# q <- q + csstyle::theme_fhi_lines_horizontal()
# q <- q + csstyle::set_x_axis_vertical()
# q
```

## Epicurves for multiple geographical areas

When we have multiple geographical areas, we use the function `lemon::facet_rep_wrap` to create multiple epicurves.

Daily epicurve for all geographical areas with vertical x-axis labels

```{r fig.height=10, fig.width=8}
# q <- ggplot(d, aes(x = date, y = N))
# q <- q + geom_col(fill = csstyle::base_color, width = 0.8)
# q <- q + lemon::facet_rep_wrap(~location_code, repeat.tick.labels = "y")
# q <- q + csstyle::scale_fill_fhi("Location",palette="primary")
# q <- q + scale_x_date("Date")
# q <- q + scale_y_continuous("Number of reported cases",
#   breaks = csstyle::pretty_breaks(5),
#   expand = expansion(mult = c(0, 0.1))
# )
# q <- q + labs(title = "Epicurve from 2018-01-01 to 2018-02-20")
# q <- q + labs(caption = fhi_caption())
# q <- q + csstyle::theme_fhi_lines_horizontal()
# q <- q + csstyle::set_x_axis_vertical()
# q
```

Weekly epicurve with vertical x-axis labels

```{r fig.height=10, fig.width=8}
# q <- ggplot(d, aes(x = isoyearweek, y = N))
# q <- q + geom_col(fill = csstyle::base_color, width = 0.8)
# q <- q + lemon::facet_rep_wrap(~location_code, repeat.tick.labels = "y", ncol=4)
# q <- q + scale_fill_fhi("Location",palette="primary")
# q <- q + scale_x_discrete("Isoweek")
# q <- q + scale_y_continuous("Number of reported cases",
#   breaks = csstyle::pretty_breaks(5),
#   expand = expansion(mult = c(0, 0.1))
# )
# q <- q + labs(title = "Epicurve from 2018-01-01 to 2018-02-20")
# q <- q + labs(caption = fhi_caption())
# q <- q + theme_fhi_lines_horizontal()
# q <- q + csstyle::set_x_axis_vertical()
# q
```

## Coloured epicurves

Sometimes you would like to add colours to differentiate between different variables. This can be done through the `fill` attribute.

```{r}
q <- ggplot(w[location_code %in% c(
  "county_nor03",
  "county_nor11",
  "county_nor15",
  "county_nor30",
  "county_nor34")], 
  aes(x = isoyearweek, y = N, fill = location_code))
q <- q + geom_col(width = 0.8)
q <- q + csstyle::scale_fill_fhi("Location",palette="primary")
q <- q + scale_x_discrete("Isoweek")
q <- q + scale_y_continuous("Number of reported cases",
  breaks = csstyle::pretty_breaks(5),
  expand = expansion(mult = c(0, 0.1))
)
q <- q + labs(title = "Epicurve from 2018-01-01 to 2018-02-20")
q <- q + labs(caption = fhi_caption())
q <- q + csstyle::theme_fhi_lines_horizontal()
q <- q + csstyle::set_x_axis_vertical()
q
```

## Right axis on epicurve

Sometimes you would like to add colours to differentiate between different variables. This can be done through the `fill` attribute.

```{r}
pd <- w[location_code %in% "county_nor03"]
pd[, cum_n := cumsum(N)]

pd[, value_left := N]
pd[, value_right := cum_n]

max_left <- max(pd$value_left)
max_right <- max(pd$value_right)

# increase the space in the y-axis
max_left <- max(c(max_left, 5))
max_right <- max(c(max_right, 5))

pd[, scaled_value_right := value_right / max_right * max_left]

q <- ggplot(pd, aes(x = isoyearweek))
q <- q + geom_col(
  mapping = aes(
    y = value_left,
    fill = "The fill variable",
  ),
  width = 0.8
  )
q <- q + geom_line(
  mapping = aes(
    y = scaled_value_right,
    group = 1,
    color = "The color variable"
  ),
  lwd = 2
  )
q <- q + scale_x_discrete("Isoweek")
q <- q + scale_y_continuous("Number of reported cases",
  breaks = csstyle::pretty_breaks(5),
  expand = expansion(mult = c(0, 0.1)),
  sec.axis = sec_axis(
    name = "Cumulative number of reported cases\n",
    ~ . * max_right / max_left,
    breaks = csstyle::pretty_breaks(5),
    labels = csstyle::format_nor
  )
)
q <- q + csstyle::scale_fill_fhi("Fill", palette = "primary")
q <- q + csstyle::scale_color_fhi("Color", palette = "posneg")
q <- q + labs(title = "Epicurve from 2018-01-01 to 2018-02-20")
q <- q + labs(caption = fhi_caption())
q <- q + csstyle::theme_fhi_lines_horizontal()
q <- q + csstyle::set_x_axis_vertical()
q
```
